Parallel Sorting of Roughly-Sorted Sequences

### 2016-12-09 tex/pres.tex
adding slides, graphics

% BEGIN CUDA INTRO
\begin{frame}
\frametitle{}
\end{frame}

\begin{frame}
\frametitle{}
\end{frame}
% END CUDA INTRO

% BEGIN PAR RADIUS
\begin{frame}
\frametitle{}
\end{frame}

\begin{frame}
\frametitle{}
\end{frame}
% END PAR RADIUS

% BEGIN PAR RADIUS RESULTS
\begin{frame}
\frametitle{}
\end{frame}

\begin{frame}
\frametitle{}
\end{frame}
% END PAR RADIUS RESULTS

@@

% BEGIN EXPLANATION OF BAD PAR. PERFORMANCE
\begin{frame}
\frametitle{}
\end{frame}

\begin{frame}
\frametitle{}
\end{frame}
% END EXPLANATION OF BAD PAR. PERFORMANCE

\begin{frame}
\frametitle{Conclusion and Further Research}
\end{frame}

\end{document}

>>>>

% BEGIN CUDA INTRO

\begin{frame}
\frametitle{Parallelization}
Roughsort may be parallelized, but assumes underlying CRCW PRAM to achieve this.
\newline\newline
CRCW PRAM:
\begin{enumerate}
\item Model for RAM which essentially says multiple slots in memory may be updated or concurrently
\item This update is done by a fixed number of processors
\item Several other restrictions to the model
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{CUDA: Our Answer for CRCW PRAM}
However, CRCW PRAM is just a useful abstraction.  The real world is just that: real
\newline\newline
Must find a proxy for CRCW PRAM to develop an implementation.
\end{frame}

\begin{frame}
\frametitle{CUDA: Our Answer for CRCW PRAM (2)}
Enter CUDA.  Similar characteristics:
\begin{enumerate}
\item Multiple slots in global memory may be updated simultaneously
\item Multiple slots may also be read simultaneously
\item Thousands more processor cores than standard CPUs, providing a mechanism to update all this parallel memory
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{CUDA: Background}
Nvidia provides APIs and an SDK to allow developers the use of GPU functions.
\newline\newline
Model:
\begin{enumerate}
	\item CPU-homed (host) program is specifically programmed to talk to GPU
	\item Host program is compiled ahead of time with CUDA hooks and calls embedded in its C++ source
	\item Payload data is explicitly shipped off to the GPU
	\item Host ships the compiled GPU code (kernel) to the GPU and launches it asynchronously
	\item GPU executes payload, places data in defined locations, host program picks up result data
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{CUDA: Background(2)}
Organization
\begin{enumerate}
	\item Core component and lowest-level work unit of CUDA execution context is a thread
	\item Up to thousands of threads may be organized in a group called a block
	\item Multiple blocks make up a grid
	\item Threads per block and other parameters are specified by launching program
	\item CUDA cheaply switches threads in and out of processing on cores
\end{enumerate}
\end{frame}

\begin{frame}
	\frametitle{CUDA: Background(3)}
	Warps
	\begin{enumerate}
		\item Groups of 32 threads which execute together are called a warp - SIMT
		\item This is a concept that is abstracted away from programmer during development/kernel launch (32 is fixed)
		\item Threads in a warp executing on a group of 32 cores are all executing the same instruction at the same time, optimally
		\item In cases of a conditional, some cores may pause to allow threads with a true value to process, then false values process - Warp divergence
	\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{CUDA: Pattern}

Threads, blocks, and grid all have identifiers
\newline\newline
Developer references these identfiers in code
\newline\newline
Using an understanding of how work is to be divided up among threads, these identifiers can inform a thread what work it needs to do

\end{frame}

\begin{frame}
	\frametitle{CUDA: Visual}
	\includegraphics[height=.9\textheight]{./cudaorg.png}
\end{frame}
% END CUDA INTRO

% BEGIN PAR RADIUS
\begin{frame}
\frametitle{Parallel Radius Finding: Translating}
Back to parallel Roughsort: we know we can find the radius in parallel.  Merely need to convert the pseudocode to CUDA code.
\newline\newline
Straightforward for $LR$ list, slightly less so for $RL$, least straightforward for $DM$.
\newline\newline
Eventually settled on a $DM$ list determination based on the sequential, with array indicies fixed to thread ID.
\newline\newline
Added an efficient method to check for full array sortness as part of the $RL$ and $LR$ list components.
\end{frame}

\begin{frame}
	\frametitle{Parallel Radius Finding: Scaling Up}
Testing:
\begin{enumerate}
	\item Worked fine for a low values
	\item Did not function at all for large values of $n$ (widely inaccurate $k$ values)
\end{enumerate}
\end{frame}

\begin{frame}
	\frametitle{Parallel Radius Finding: Re-Visiting}
	Cause/Fix:
	\begin{enumerate}
		\item Asked individual threads to allocate their slot in the global memory $RL$ and $LR$ lists upon startup
		\item Not all threads are immediately scheduled for large thread counts--bug
		\item Subsequently started copying all values before kernel launch
	\end{enumerate}

\end{frame}

\begin{frame}
	\frametitle{Parallel Radius Finding: Scaling Up (2)}
	After fixing, $DM$ list was correct to an extent, but it was slow.
	\newline\newline
	Resolved this by realizing only need our "radius" to approximate the real $k$ value, in that we just need the next power of two greater
	\newline\newline
	Added \texttt{Thrust max_element} code to find the greatst value in the $DM$ list, thus yielding our radius
\end{frame}
% END PAR RADIUS

% BEGIN PAR RADIUS RESULTS
\begin{frame}
\frametitle{Parallel Radius Finding: Results ($k = 2$)}
\begin{figure}
\resizebox{!}{.8\textheight}{\input{./plots/seqpar2.tex}}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Parallel Radius Finding: Results($k = 100$)}
\begin{figure}
	\resizebox{!}{.8\textheight}{\input{./plots/seqpar100.tex}}
\end{figure}
\end{frame}
% END PAR RADIUS RESULTS

@@

% BEGIN EXPLANATION OF BAD PAR. PERFORMANCE
\begin{frame}
\frametitle{Parallel Performance}
Below expectations:
\begin{enumerate}
	\item Parallel radius runs in linear time
	\item Parallel Roughsort by Mergesort consistently
\end{enumerate}
Why: CUDA is not CRCW PRAM
\begin{enumerate}
	\item Warp divergence
	\item Memory coalescence
\end{enumerate}
\end{frame}
% END EXPLANATION OF BAD PAR. PERFORMANCE

\begin{frame}
\frametitle{Conclusion and Further Research}
Demonstrated a good method for evaluating sorting algorithms, by being able to generate $k$-sorted sequences
\newline\newline
Developed a first implementation of Roughsort\textemdash to be continued by others
\newline\newline
Bottlenecks in implementation should be researched so they may be resolved
\end{frame}
\end{document}
